---
title: "Final Project on Framingham"
author: "Hunter Manhart and Arturo Perez"
date: "December 5, 2017"
output: html_document
---

##Load library and data
We need the glmnet package
```{r, eval=FALSE}
install.packages("glmnet", repos = "http://cran.us.r-project.org")
```

Then load it in
```{r, message=FALSE, warning=FALSE}
library(glmnet)
```

Now get the Framingham data (make sure frm.csv is in the same directory)
```{r, eval=T}
framingham <- read.csv("frm.csv")
```

Let's glance at how the data is organized
```{r, eval=T}
str(framingham)
```


##Divide and scale data
Get the titles of the columns of data
```{r}
fnames=names(framingham)
fnames
```

Split into factors and events
```{r}
FactorNames=fnames[seq(1,24)]
EventNames=fnames[seq(25,40)]
```

Scale the factored data
```{r}
fhscaled<-framingham
fhscaled[FactorNames]<-as.data.frame(scale(framingham[FactorNames]))
str(fhscaled)
```


##Mortality

How common was death
```{r}
sum(framingham['DEATH'])
```

Let's isolate the data of those who died
```{r, eval=T}
deaths<-subset(framingham, DEATH==1)
eventTimes<-c('TIMEMI','TIMEMIFC','TIMECHD','TIMESTRK','TIMECVD','TIMEDTH','TIMEHYP')
plot(deaths[eventTimes])
```
```{r}
events<-setdiff(eventTimes,cbind("TIMEDTH"))
x<-as.matrix(deaths[events])
y<-as.matrix(deaths$TIMEDTH)
str(x)
str(y)
fit <- cv.glmnet(x, y, alpha=1)  #alpha=1 is lasso, alpha=0 is ridge, lambda =0 is linear regression
plot(fit)
```
```{r}
beta<-coef(fit,s=2.718^-4)
beta
```


##Isolate to Third Period
```{r}
nacols<-apply(framingham, 2, function(x) {sum(is.na(x))})
nacols
```

```{r}
fnamesKeep <- setdiff(fnames,cbind("HDLC", "LDLC","GLUCOSE"))
fhs.noHLG<-fhscaled[fnamesKeep]
fhs.noHLG<-fhs.noHLG[complete.cases(fhs.noHLG),]
str(fhs.noHLG)
```

```{r}
thirdPeriod <-subset(fhscaled, PERIOD>1)
thirdPeriod <- thirdPeriod[complete.cases(thirdPeriod),]
str(thirdPeriod)
```

##Logistic Regression
```{r}
x <- as.matrix(thirdPeriod[FactorNames])
y <- as.matrix(thirdPeriod$HYPERTEN)

fit = glmnet(x, y, family = "binomial")
plot(fit)

vnat=coef(fit)
vnat=vnat[-1,ncol(vnat)] # remove the intercept, and get the coefficients at the end of the path
axis(4, at=vnat,line=-.5,label=colnames(x),las=1,tick=FALSE, cex.axis=0.5)
```


```{r}
fkeep <- setdiff(FactorNames, 'DIABETES')
x <- as.matrix(thirdPeriod[fkeep])

diabetes <-subset(framingham, PERIOD==3)
diabetes <- diabetes[complete.cases(diabetes),]
y <- as.matrix(diabetes$DIABETES)
fit = glmnet(x, y, family = "binomial")
plot(fit)

vnat=coef(fit)
vnat=vnat[-1,ncol(vnat)] # remove the intercept, and get the coefficients at the end of the path
axis(4, at=vnat,line=-.5,label=colnames(x),las=1,tick=FALSE, cex.axis=0.5)
```

```{r}
beta<-coef(fit,s=2.718^-6)
beta
```

##Compare to linear regression
```{r}
linear = glmnet(x,y, alpha=0)
plot(linear)

vnat=coef(linear)
vnat=vnat[-1,ncol(vnat)] # remove the intercept, and get the coefficients at the end of the path
axis(4, at=vnat,line=-.5,label=colnames(x),las=1,tick=FALSE, cex.axis=0.5)
```